Microbial interactions as drivers of a nitrifying ecosystem in a chemostat

### 2020-03-24 tex/Article_Interactions.tex
Changes in .tex and associated graphics

Consider the case where $n_1 = 1$ and $n_2 = 1$. Two set of parameters for For gaining some insight the comparison of the operating diagrams where no interactions take place ($A = 0$) and with a non-zero, as shown in \ref{case1_no_interactions} and \ref{case1_interactions}. When $A=0$ the analysis above presented is no longer valid (A is not invertible), nevertheless the stability analysis is much simpler and is given in the supplementary material section. Parameters can be seen in table \ref{kinetic_parameters_case_study_1}, while the interaction matrices.

@@

Kinetic parameters of case study 2 can be seen in table \ref{kinetic_parameters}. Interaction matrices of each case are shown in figure \ref{Interactions}. Case 1 was taken from Dumont \textit{et al.}\cite{Dumont2016} where they proposed a distribution of parameters obtained with a Bayesian estimation method. Their fit describes well the dynamics of the two most abundant OTU in each functional group, but it still fails to capture the measured substrates dynamics.

@@

\begin{figure}[h]
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{Stability/Interactions_parameters_Dumont}
	\caption{Originally calibrated interaction matrix.}
	\label{Interaction_1}
\end{subfigure}
~
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{Stability/Interactions_parameters_modified}
	\caption{Modified interaction matrix.}
	\label{Interaction_2}
\end{subfigure}
	\caption{Interaction matrices for each case for a consortia of 4 bacterial species where $x_1$ and $x_2$ are AOB and $x_3$ and $x_4$ are NOB. Only the entries $a_{11}$, $a_{12}$, and $a_{21}$ were modified, accounting for reciprocal positive interactions between $x_1$ and $x_2$.}
	\label{Interactions}
\end{figure}

The ecological diagrams for each case are presented in figure \ref{ecological_diagrams}. One can see that coexistence in the same functional group is never attained for case 1, whereas for case 2 a more complex situation takes places; even a zone where more than one stable equilibria is found, implying that the outcome in that zone depends on initial conditions of the system. Note that CN can be achieved at a more elevated dilution rate in case 2 than in case 1, which can be explained by noting that the interactions of $x_1$ on $x_3$ and $x_4$ are positive and stronger in case 2, thus helping them survive through bigger dilution rates.

 Ecological diagrams also show that the model is able to overcome the competitive exclusion principle \cite{lobry2017chemostat} (CEP) paradigm. The CEP roughly states that if two species are growing on the same limiting resource, and their growth laws are non-decreasing in the resource, then only one of them will survive in the long run. However reports on wastewater treatment plants have shown coexistence between species in nitrifying reactors \cite{Wagner2002}, thus implying that a more complex growth law (as shown in here) or model structure involving other biological processes is required to include microbial diversity.

@@

\begin{figure}[ht]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{Stability/ED_parameters_Dumont}
		\caption{Original parameters ecological diagram.}
	\end{subfigure}
~
	\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{Stability/ED_parameters_modified}
	\caption{Modified interaction matrix ecological diagram.}
	\end{subfigure}
	\caption{Ecological diagrams for both cases. The different zones represent the combination of surviving species in the steady state. PN takes place when neither $x_3$ nor $x_4$ are present. MSE means multiple stable equilibria. CN takes place if $x_3$ or $x_4$ are present. A complex situation emerges in ecological terms for case 2, depending on the operating parameters the community can go from a single species surviving in the chemostat to full coexistence. }
	\label{ecological_diagrams}
\end{figure}

%
%\begin{figure}[ht]
%	\centering	
%	\begin{subfigure}[b]{0.45\textwidth}
%		\includegraphics[width=\textwidth]{Stability/OD_parameters_Dumont}
%		\caption{Case 1 operating diagram.}
%	\end{subfigure}
%~
%	\begin{subfigure}[b]{0.45\textwidth}
%		\includegraphics[width=\textwidth]{Stability/OD_parameters_modified}
%		\caption{Case 2 operating diagram.}
%	\end{subfigure}
%	\caption{Operating Diagrams for both cases. Case 1 Operating diagram coincides with the ecological diagram. Note how in case 2 CN might occur at bigger dilution rates than case 1. There is also a thin line where both CN and PN may arise depending on initial conditions.}
%	\label{operating_diagrams}
%\end{figure}
%
%

>>>>

Consider the case where $n_1 = 1$ and $n_2 = 1$. For gaining some insight the comparison of the operating diagrams where no interactions take place ($A = 0$) and with a non-zero, as shown in \ref{case1_no_interactions} and \ref{case1_interactions}. When $A=0$ the analysis above presented is no longer valid (A is not invertible), nevertheless the stability analysis is much simpler and is given in the supplementary material section. Parameters can be seen in table \ref{kinetic_parameters_case_study_1}, while the interaction matrices can be seen .

@@

 The case study 2 is based on Dumont \textit{et al.}\cite{Dumont2016} model parameters. Again for studying the model behaviour a modification of the original interaction matrix is performed. They proposed a distribution of parameters obtained from a Bayesian estimation method. Their fit describes well the dynamics of the two most abundant OTU in each functional group, but it still fails to capture the measured substrates dynamics. Kinetic parameters of case study 2 can be seen in table \ref{kinetic_parameters}. The original interaction matrix estimated by them is seen in \ref{Interaction_1}. A second matrix is presented, which is obtained from modification on the sign of coefficient $a_{11}$ as seen in \ref{Interaction_2}, and finally a third one is obtained by changing the value of $a_{12}$ from the original matrix.

@@

\begin{figure}[h]
\begin{subfigure}[b]{0.3\textwidth}
	\includegraphics[width=\textwidth]{Stability/Interactions_parameters_Dumont}
	\caption{Originally calibrated interaction matrix.}
	\label{Interaction_1}
\end{subfigure}
~
\begin{subfigure}[b]{0.3\textwidth}
	\includegraphics[width=\textwidth]{Stability/Interactions_parameters_modified}
	\caption{Modified interaction matrix with positive intra species interaction for $x_1$.}
	\label{Interaction_2}
\end{subfigure}
~
\begin{subfigure}[b]{0.3\textwidth}
	\includegraphics[width=\textwidth]{Stability/Interactions_parameters_modified_2}
	\caption{Modified interaction matrix with positive intra species interaction for $x_1$.}
	\label{Interaction_3}
\end{subfigure}
	\caption{Interaction matrices for each case for a consortia of 4 bacterial species where $x_1$ and $x_2$ are AOB and $x_3$ and $x_4$ are NOB. Only the sign of the entry $a_{12}$ was modified with respect to the original matrix. Representing the a positive interspecies interaction of $x_2$ on $x_1$.}
	\label{Interactions}
\end{figure}

@@

\begin{figure}[ht]
	\centering
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{Stability/ED_parameters_Dumont}
		\caption{Original parameters ecological diagram.}
	\end{subfigure}
~
	\begin{subfigure}[b]{0.3\textwidth}
	\includegraphics[width=\textwidth]{Stability/ED_parameters_modified}
	\caption{Modified interaction matrix ecological diagram.}
	\end{subfigure}
~
\begin{subfigure}[b]{0.3\textwidth}
	\includegraphics[width=\textwidth]{Stability/ED_parameters_modified_2}
	\caption{Modified interaction matrix ecological diagram.}
\end{subfigure}

	\caption{Ecological diagrams for both cases. The different zones represent the combination of surviving species in the steady state. PN takes place when neither $x_3$ nor $x_4$ are present. MSE means multiple stable equilibria. CN takes place if $x_3$ or $x_4$ are present. A complex situation emerges in ecological terms for case 2, depending on the operating parameters the community can go from a single species surviving in the chemostat to full coexistence. }
	\label{ecological_diagrams}
\end{figure}

The ecological diagrams for each case are presented in figure \ref{ecological_diagrams}. One can see that coexistence in the same functional group is never attained for case 1, whereas for case 2 $x_1$ and $x_2$, both AOB, coexist in either partial or complete nitrification. Ecological diagrams also show that the model is able to overcome the competitive exclusion principle \cite{lobry2017chemostat} (CEP) paradigm. The CEP roughly states that if two species are growing on the same limiting resource, and their growth laws are non-decreasing in the resource, then only one of them will survive in the long run. However reports on wastewater treatment plants have shown coexistence between species in nitrifying reactors \cite{Wagner2002}, thus implying that a more complex growth law (as shown in here) or model structure involving other biological processes is required to include microbial diversity.

\clearpage
%
%\begin{figure}[ht]
%	\centering	
%	\begin{subfigure}[b]{0.45\textwidth}
%		\includegraphics[width=\textwidth]{Stability/OD_parameters_Dumont}
%		\caption{Case 1 operating diagram.}
%	\end{subfigure}
%~
%	\begin{subfigure}[b]{0.45\textwidth}
%		\includegraphics[width=\textwidth]{Stability/OD_parameters_modified}
%		\caption{Case 2 operating diagram.}
%	\end{subfigure}
%	\caption{Operating Diagrams for both cases. Case 1 Operating diagram coincides with the ecological diagram. Note how in case 2 CN might occur at bigger dilution rates than case 1. There is also a thin line where both CN and PN may arise depending on initial conditions.}
%	\label{operating_diagrams}
%\end{figure}
%
%
